---
title: "Insert Title"
author: |
  | Author: Holiday Tang 
  | [LinkeDin](https://www.linkedin.com/in/holiday-t/)
  | [GitHub](https://github.com/holiday007) | [Kaggle Novice](https://www.kaggle.com/holibae007)
date: "Date: `r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
    toc: true
    fig_width: 7
    fig_height: 5
    toc_depth: 2
    dev: png
---

Themes: cayman, architect, leonids, hpstr

### This template does not support code-folding

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, comment=NA, cache=T)
```

```{r}
3+3
```

### Maybe some of the github options will work but I don't know

Use includes option to include extra content, for example

```
includes:
  in_header: header.md
  before_body: doc_prefix.md
  after_body: something.md
```

Markdown Extension:

* By default R Markdown is defined as all pandoc markdown extensions with the following tweaks for backward compatibility with the markdown package:

```
+autolink_bare_uris
+ascii_identifier
+tex_math_single_backslash
```

* You can enable or disable markdown extensions using the md_extensions option (you preface an option with - to disable and + to enable it). For example:

```
---
title: "Habits"
output:
  github_document:
    md_extensions: -autolink_bare_uris+hard_line_breaks
---
```

* also allows integration of pandoc features through option, pandoc_args:


## Shared Options:

If you want to specify a set of default options to be shared by multiple documents within a directory you can include a file named _output.yaml within the directory. Note that no YAML delimiters or enclosing output object are used in this file. For example:

_output.yaml

```
github_document:
  fig_width: 5
  fig_height: 5
```


