---
title: "Natural Language Processing Practice"
author: |
  | Author: Holiday Tang
  | [LinkeDin](https://www.linkedin.com/in/holiday-t/) |  [GitHub](https://github.com/holiday007) | [Kaggle Novice](https://www.kaggle.com/holibae007)
date: "Date: `r Sys.Date()`"
output:
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
    toc_depth: 2
    dev: png
    pandoc_args: --webtex
---

(Note: adopted from *Data Scicence and Machine Learning Bootcamp with R* by *Jose Portilla* from Udemy) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, comment=NA, cache=T)

# install packages
packs = c()

new.pkg <- packs[!(packs %in% installed.packages())]

if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}
```

# Introduction to natural language processing

Potential purpose:
  
  - Find relevant documents
  - Group articles by topic
  
We will want to:

* compile documents
* Featurize them
* Compare their features

Featurization:
  
  - one option is based on **word counts**
  - use consine similarity on the vectors made to determine similarity:

$$
sim(A,B) = cos(\theta) = \frac{A\cdot B}{||A||*||B||}
$$

We can improve on Bag of words by adjusting word counts based on their frequency in corpus, the group of all documents.

We can use TF-IDF (Term Frequency-Inverse Document Frequency) 

* Term frequency - **Importance** of the term within that document
  - TF(d,t) = Number of occurences of term t in document d

* Inverse document frequency - how important a word is to a document in a collection or corpus. 
  - IDF(t) = log(D/t) 
    - D = total number of documents 
    - t = number of documents with the term
    - so for term x within document y

$$
W_{x,y} = tf_{x,y} \cdot log(N/df_x)
$$

- $tf_{x,y}$ - frequency of x in y
- $df_x$ - number of documents containing x
- N - total number of documents

The tf–idf value increases proportionally to the number of times a word appears in the document and is offset by the number of documents in the corpus that contain the word, which helps to adjust for the fact that some words appear more frequently in general.

```{r}

```


...package

""