---
title: "Insert Title"
author: |
  | Author: Holiday Tang 
  | [LinkeDin](https://www.linkedin.com/in/holiday-t/) |  [GitHub](https://github.com/holiday007) | [Kaggle Novice](https://www.kaggle.com/holibae007)
date: "| Date: `r Sys.Date()`"
output:
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
    toc_depth: 2
    dev: png
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, comment=NA, cache=T)

# install packages if necessary
packs = c("ggplot2", "rpart", "randomForest", "caTools", "rpart.plot")

new.pkg <- packs[!(packs %in% installed.packages())]

if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}
```

# Tree Methods Project

##### Data

```{r}
College = ISLR::College
head(College)
```

```{r}
# uncomment following line for data set description
# ?ISLR::College
```

#### EDA

**Explore data a little bit**

Scatterplot of Grad.Rate versus Room.Board, colored by the Private column

```{r}
library(ggplot2)

ggplot(College, aes(x=Room.Board, y=Grad.Rate, color=Private)) +
  geom_point()
```

Histogram of graduation rate by private

```{r}
ggplot(College, aes(Grad.Rate)) + 
  geom_histogram(aes(fill = Private), color = "black", bins = 50)
```

There is a college with an graduation rate above 100%, changing is to 100

```{r}
College[which(College$Grad.Rate > 100),]
College[which(College$Grad.Rate > 100), 18] = 100
```

#### Train Test Split

```{r}
library(caTools)

Split = sample.split(College$Private, SplitRatio = 0.7)
train.data = subset(College, Split == T)
test.data = subset(College, Split == F)

# head(train.data)
```

#### Decision Tree

Building model

```{r}
library(rpart)
tree = rpart(Private ~ ., method = "class", data = train.data)

# predict
predicted.private = predict(tree, test.data)
head(predicted.private)
```

```{r}
predicted.private = data.frame(predicted.private)
predicted.private$Private = predicted.private$Yes > predicted.private$No

library(dplyr)
predicted.private = predicted.private %>%
  mutate(Private = ifelse(Private == T, "Yes", "No"))

head(predicted.private)
```

Confusion Matrix
```{r}
tree.confusion = table(predicted.private$Private, test.data$Private)
tree.confusion
```

Error rate
```{r}
(tree.confusion[1,2] + tree.confusion[2,1])/sum(tree.confusion)
```

Plot the tree
```{r}
library(rpart.plot)
prp(tree, main = "Private or not")
```

#### Random Forest

```{r}
library(randomForest)

rf.model = randomForest(Private ~ ., importance = T, data = train.data)

rf.model
```

Model confusion matrix

```{r}
rf.model$confusion
```

```{r}
rf.model$importance
```

Prediction

```{r}
rf.predict = predict(rf.model, test.data)

rf.confusion = table(rf.predict, test.data$Private)

rf.confusion
```

Error rate:

```{r}
(rf.confusion[1,2] + rf.confusion[2,1])/sum(rf.confusion)
```

