---
title: "ggplot2"
author: |
  | Author: Holiday Tang 
  | [LinkeDin](https://www.linkedin.com/in/holiday-t/) |  [GitHub](https://github.com/holiday007) | [Kaggle Novice](https://www.kaggle.com/holibae007)
date: "| Date: `r Sys.Date()`"
output:
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
    toc_depth: 3
    dev: png
---

### Replicating exercise from 

[] (https://cedricscherer.netlify.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/)

```{r, message=FALSE, warning=FALSE, eval=FALSE}
packs = c("ggplot2", "ggthemes", "tidyverse", "extrafont", 
          "cowplot", "grid", "gridExtra", "ggrepel", 
          "reshape2", "ggforce", "ggridges", "shiny", "patchwork",
          "readr", "tidyverse", "extrafont")

for (i in length(packs)){
  if(!require(packs[i], quietly = TRUE)) install.packages(packs[i])
}
```

```{r, message = FALSE}
chic = readr::read_csv("https://raw.githubusercontent.com/Z3tt/R-Tutorials/master/ggplot2/chicago-nmmaps.csv")

tibble::glimpse(chic)
head(chic, 10)
```

### elements of ggplot2

* Data
* Geometries - `geom_`: geometric shape that represents the data
* Aesthetics - `aes()`: Aesthetics of the geometric and statistical objects, such as color, size, shape, transparency and position
* Scales `scale_`: Maps between the data and the aesthetic dimensions, such as data range to plot width or factor values to colors.
* Statistical transformations `stat_`: Statistical summaries of the data, such as quantiles, fitted curves and sums.
* Coordinate system `coord_`: The transformation used for mapping data coordinates into the plane of the data rectangle.
* Facets `facet_`: The arrangement of the data into a grid of plots.
* Visual themes `theme()`: The overall visual defaults of a plot, such as background, grids, axes, default typeface, sizes and colors.

```{r}
# library(ggplot2)
library(tidyverse)
```

```{r}
# start by defining a plot element
# ggplot(data = df, aes(x = , y = ,))

(g = ggplot(chic, aes(x=date, y=temp))) 
#using parentheses while creating an object the object will be printed immediately 
```

#### Scatterplot

```{r}
g + geom_point() # scatterplot
```

```{r}
g + geom_point(color = 'purple')
```

##### Setting theme

```{r}
theme_set(theme_bw())

g + geom_point(color = 'purple')
```

### Working with Axes

#### Add Axis Labels

* Temperature - degree ~ F 

```{r}
ggplot(chic, aes(x=date, y=temp)) +
  geom_point(color='purple') + 
  labs(x='Year', y=expression(paste('Temperature (', degree ~ F, ')')))
```

#### Manipulating Labels

* `theme()` - all kinds of elements ()[https://ggplot2.tidyverse.org/reference/theme.html]

```{r}
ggplot(chic, aes(x=date, y=temp)) +
  geom_point(color = 'purple') + 
  labs(x='Year', y='Temperature (°F)') + 
  theme(axis.title.x = element_text(color = 'red', size = 15, vjust = 1),
        axis.title.y = element_text(color = 'blue', size = 15, vjust = 1))
```

* `angle`, `vjust` - adjust position of the text

```{r}
ggplot(chic, aes(x=date, y=temp)) +
  geom_point(color='purple') +
  labs(x="Year", y="Temperature (°F)") + 
  theme(axis.text.x = element_text(angle = 45, size = 15, vjust = 0.5))
```

#### Remove Axis Ticks & Tick Text

*`element_blank()` - magic of getting rid of something

```{r}
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(color="purple") + 
  labs(x="Year", y="Temperature (°F)") + 
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```

#### Limit Axis Range

```{r}
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(color="purple") + 
  labs(x="Year", y="Temperature (°F)") +
  ylim(c(0,50))
```

Alternative:

* `g + scale_x_continuous(limits = c(0, 50))`: remove out-of-range points
* `g + coord_cartesian(xlim = c(0, 50))` : adjust visible area

#### Force plot to start at origin

```{r}
chic %>%
  dplyr::filter(temp>25, o3>20) %>%
  ggplot(aes(x=temp, y=o3)) + 
  geom_point(color = "pink") + 
  labs(x=expression(paste("Temperature higher than 25", degree ~ F, "")), y="Ozone higher than 20 ppb") + 
  expand_limits(x=0, y=0)
```

Alternative: `coord_cartesian(xlim = c(0, max(chic_red$temp)), ylim = c(0, max(chic_red$o3)))`

```{r}
ggplot(chic, aes(x=temp, y=temp + rnorm(nrow(chic), sd=20))) + 
  geom_point() + 
  xlim(c(0,100)) + ylim(c(0,150)) +
  coord_equal()
```

#### Altering Labels by Function

```{r}
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(color="purple") + 
  labs(x="Year", y="Temperature (°F)") + 
  scale_y_continuous(labels = function(x) {return(paste(x, "Degrees Fahrenheit"))})
```

### Titles

* `ggtitle()`

```{r}
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(color="purple") + 
  labs(x="Year", y="Temperature (°F)") + 
  ggtitle("Temperatures in Chicago") 
```  

* `lab(title = , subtitle = , caption = )`

```{r}
ggplot(chic, aes(x = date, y = temp)) +
  geom_point(color = "firebrick") + 
  labs(x = "Year", y = "Temperature (°F)",
       title = "Temperatures in Chicago",
       subtitle = "Seasonal patter of daily temperatures from 1997 to 2001",
       caption = "Data:NMAPS"
  )
```


#### Title Manipulation

* Make Title Bold and Add Space at Baseline

```{r}
ggplot(chic, aes(x = date, y = temp)) +
  geom_point(color = "firebrick") + 
  labs(x = "Year", y = "Temperature (°F)", 
       title = "Temperatures in Chicago") +
  theme(plot.title = element_text(size=15, face="bold", 
                                  margin = margin(10,0,8,0)))
```

##### Position of Title

```{r}
ggplot(chic, aes(x = date, y = temp)) +
  geom_point(color = "firebrick") +  
  labs(x = "Year", y = "Temperature (°F)", 
       title = "Temperatures in Chicago") + 
  theme(plot.title = element_text(size = 15, face = 9, hjust = 0.5))
```

#### Fonts

* `extrafont::font_import()`  
* `element_text(..., family = ...)`

```{r}
library(extrafont)
font_import()
```

```{r}
fonts()
```

```{r}
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(color="purple") + 
  labs(x="Year", y="Temperature (°F)", title="Temperatures in Chicago") + 
  theme(plot.title = element_text(size=18, family = "Times New Roman"))
```

```{r}
# setting default fonts
theme_set(theme_bw(base_size = 14, base_family = "Calibri"))
```

#### Spacing in multi-line text

```{r}
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(color = "purple") + 
  labs(x="Year", y="Temperature (°F)") + 
  ggtitle("Temperatures in Chicago\nfrom 1997 to 2001") + # notice \n
  theme(plot.title = element_text(size=16, face="bold", vjust=1,
                                  lineheight = 1))
```

#### Legends

```{r}
# color code based on factor
ggplot(chic, aes(x=date, y=temp)) + 
  geom_point(aes(color = factor(season))) +  # to access variable you got to
  # put it in aes function
  labs(x="Year", y="Temperature (°F)")
```

- `theme(legend.position = "none")`
- `guides(fill = F)`
- `scale_fill_discrete(guide = F)`

```{r}
# turn off legend in theme (function)
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  theme(legend.position = "none")
```


```{r}
# turn off legen title
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  theme(legend.title = element_blank())
```

```{r}
# legend position - top, right, bottom, left
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  theme(legend.position = "top", legend.title = element_blank())
```

```{r}
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  theme(legend.title = element_text(size=14, color="orange",
                                    face="bold"))
```

```{r}
# change legend title
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() + 
  labs(x="Year", y="Temperature (°F)", color="Seasons:") 
```

legend details can be changed by:

* `scale_color_discrete`  
* `scale_color_continuous`

```{r}
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  theme(legend.title = element_text(color = "chocolate", 
                                    size = 14, face = "bold")) +
  scale_color_discrete(name = "Season")
```

```{r}
# change order of legend keys

#change the levels of season
chic$season = factor(chic$season, levels = c("Spring", "Summer",
                                             "Autumn", "Winter"))

ggplot(chic, aes(x=date, y=temp)) +
  geom_point(aes(color=season)) + 
  labs(x = "Year", y = "Temperature (°F)")
```

```{r}
# change legend labels
ggplot(chic, aes(x = date, y = temp, color = season)) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)", color = "Seasons:") + 
  theme(legend.title = element_text(color = "chocolate", 
                                    size = 14, face = 2)) + 
  scale_color_discrete(labels = c("Mar - May", "Jun - Aug", 
                                              "Sep - Nov", "Dec - Feb"))
```

```{r}
# Change Background Boxes in the Legend

ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  
  
  theme(legend.key = element_rect(fill = NA))
```

- `legend.key(fill = ...)`

```{r}
# Change Size of Legend Symbols
ggplot(chic, aes(x = date, y = temp, color = factor(season))) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") + 
  theme(legend.key = element_rect(fill = NA),
        legend.title = element_text(color = "chocolate", 
                                    size = 14, face = 2)) +
  scale_color_discrete("Seasons:") +
  
  
  guides(color = guide_legend(override.aes = list(size=3)))
```

* Let’s say you have a point layer and you add a rug plot of the same data. By default, both the points and the “line” end up in the legend like this:

```{r}
# leave a layer off the legend
```

