---
title: "Support Vector Machine Small Project"
author: |
  | Author: Holiday Tang 
  | [LinkeDin](https://www.linkedin.com/in/holiday-t/) |  [GitHub](https://github.com/holiday007) | [Kaggle Novice](https://www.kaggle.com/holibae007)
date: "| Date: `r Sys.Date()`"
output:
  github_document:
    toc: true
    fig_width: 7
    fig_height: 5
    toc_depth: 2
    dev: png
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, comment=NA, cache=T)

# install packages
packs = c("ggplot2", "caTools", "e1071")

new.pkg <- packs[!(packs %in% installed.packages())]

if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}


```

# Project Overview Leading Club 2016

Leading Club connect investors and investees 

Data contains a series of variables, one of which is whether the borrower paid back their loan in full.

**Project Goal**: Build a model that predict and classify whether a borrow paid back in full.

Variables: 

* credit.policy: 1 if the customer meets the credit underwriting criteria of LendingClub.com, and 0 otherwise.
* purpose: The purpose of the loan (takes values "credit_card", "debt_consolidation", "educational", "major_purchase", "small_business", and "all_other").
* int.rate: The interest rate of the loan, as a proportion (a rate of 11% would be stored as 0.11). Borrowers judged by LendingClub.com to be more risky are assigned higher interest rates.
* installment: The monthly installments owed by the borrower if the loan is funded.
* log.annual.inc: The natural log of the self-reported annual income of the borrower.
* dti: The debt-to-income ratio of the borrower (amount of debt divided by annual income).
* fico: The FICO credit score of the borrower.
* days.with.cr.line: The number of days the borrower has had a credit line.
* revol.bal: The borrower's revolving balance (amount unpaid at the end of the credit card billing cycle).
* revol.util: The borrower's revolving line utilization rate (the amount of the credit line used relative to total credit available).
* inq.last.6mths: The borrower's number of inquiries by creditors in the last 6 months.
* delinq.2yrs: The number of times the borrower had been 30+ days past due on a payment in the past 2 years.
* pub.rec: The borrower's number of derogatory public records (bankruptcy filings, tax liens, or judgments).

### Data

##### Data Overview

**Stucture**

```{r}
loans = read.csv("loan_data.csv")
str(loans)
```

**Summary**

```{r}
summary
```

Convert some variables to categorical variable

```{r}
loans$credit.policy = factor(loans$credit.policy)
loans$not.fully.paid = factor(loans$not.fully.paid)
```

### A bit of EDA

```{r}
library(ggplot2)
ggplot(loans, aes(fico)) + 
  geom_histogram(aes(fill = not.fully.paid), color = "black") + 
  ggtitle("fico score")
```

```{r}
ggplot(loans, aes(purpose)) + 
  geom_bar(aes(fill=not.fully.paid), position = "dodge", color="black") + 
  ggtitle("Purpose") + 
  theme(axis.text.x=element_text(angle = 90))
```

```{r}
ggplot(loans, aes(x=int.rate, y=fico, color=not.fully.paid)) +
  geom_point(alpha = 0.2) + 
  geom_smooth()
  ggtitle("Interest Rate vs. FICO Credit Score")
```

Higher interest rate for lower credit score, makes sense.

### Model

Train-test split

```{r}
library(caTools)

Split = sample.split(loans$not.fully.paid, SplitRatio = 0.7)
train = subset(loans, Split == T)
test = subset(loans, Split == F)
```

```{r}
library(e1071)
SVM = svm(not.fully.paid ~., data = train)
summary(SVM)
```

Prediction

```{r}
pred.values = predict(SVM, test)
table(pred.values, test$not.fully.paid)
```

Very high false negative rate. Not paid back predicted as paid back.

### Tuning parameters

cost & gamma

```{r, eval=F}
tune.result = tune(svm, train.x = not.fully.paid ~ ., data = train,
                   kernel = 'radial', 
                   ranges = list(cost=c(0.1,0.5,1,2),
                                gamma=c(0.1,0.25,0.75,1)))
```

```{r, eval=F}
tune.result
```

Takes forever to run, so I didn't include the process, but the outcome is cost = 1, and gamma = 0.75 for the values I tested

Final model:

```{r}
final.model = svm(not.fully.paid ~ ., cost = 1, gamma = 0.75, data = train)
```

Prediction by final model

```{r}
final.pred = predict(final.model, test)
table(final.pred, test$not.fully.paid)
```

Still very bad performance lol, better tune other values or use other models.

